---
title: "p8105_hw1_ns3782"
author: "NSK"
date: "2024-09-21"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1

```{r setup_data}

## Initialising project libraries and data

library(palmerpenguins)
library(tidyverse)

data("penguins", package = "palmerpenguins")

```

The penguins dataset contains data on `r nrow(penguins)` penguins, with `r ncol(penguins)` variables providing descriptive information for each penguin. These variables include species (`r unique(penguins$species)`), island (`r unique(penguins$island)`), bill length (mm), bill depth (mm), flipper length (mm), body mass (g), sex, and year (year data was collected).

The mean flipper length of the penguins is `r round(mean(penguins$flipper_length_mm, na.rm = TRUE), 2)` mm.

```{r penguins_scatterplot}

## Scatterplot of Flipper Length vs Bill Length in Penguins

ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) +
  geom_point() +
  labs(title = "Flipper Length vs Bill Length in Penguins",
       x = "Bill Length (mm)",
       y = "Flipper Length (mm)",
       color = "Species")

## Save scatterplot

ggsave("penguin_scatter.png", width = 8, height = 6)

```

## Problem 2

```{r problem_2}

## Set seed for reproducibility
set.seed(123)

## Creating dataframe with numeric, logical and character vectors

problem2_df = tibble(
  norm_samp = rnorm(10),
  vec_logical = norm_samp > 0,
  vec_char = letters[1:10],
  vec_factor = factor(c("Level1", "Level2", "Level3", "Level1", "Level2", "Level3", "Level1", "Level2", "Level3", "Level1"))
)

```

Attempting to take the mean of each variable in problem2_df dataframe.

```{r problem_2_dfmean, error=TRUE}

## Mean of numeric vector
mean_numeric = mean(pull(problem2_df, norm_samp))

## Mean of logical vector
mean_logical = mean(pull(problem2_df, vec_logical))

## Mean of character vector
mean_char = mean(pull(problem2_df, vec_char))

## Mean of factor vector
mean_factor = mean(pull(problem2_df, vec_factor))

```

We can see that the numeric and logical vectors return numeric values. In the case of the logical vector from problem2_df, this is the proportion of TRUE values as R computes TRUE as 1 and FALSE as 0.

vec_char produces an error message as it is not possible for R to calculate the mean of character values. Similarly, attempts to calculate the average directly for vec_factor does not work as factors are categorical.

Applying the as.numeric() function to the logical, character, and factor variables:

```{r problem_2_asnumeric}

## Converting logical vector to numeric
as.numeric(pull(problem2_df, vec_logical))

## Converting character vector to numeric
as.numeric(pull(problem2_df, vec_char))

## Converting factor vector to numeric
as.numeric(pull(problem2_df, vec_factor))

```

The as.numeric() function is able to convert logical variables like vec_logical successfully, as it treats TRUE as 1 and FALSE as 0. However the function is not able to work with character variables such as those in vec_char and produces NA values, as letters do not have an intrinsic integer value.

For factor variables like vec_factor, as.numeric() converts them based on their underlying integer levels, not their labels. Factors in R use underlying integer values to store categorical data, and each unique category (level) is assigned a numeric code that corresponds to its text label.
Factors can be converted to numbers, however, this is not automatically performed when calculating means using the mean() function. This is why mean() does not work on factor variables unless converted.
