---
title: "p8105_hw1_ns3782"
author: "NSK"
date: "2024-09-21"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1

```{r setup_data}

## Initialising project libraries and data

library(palmerpenguins)
library(tidyverse)

data("penguins", package = "palmerpenguins")

```

The penguins dataset contains data on `r nrow(penguins)` penguins, with `r ncol(penguins)` variables providing descriptive information for each penguin. These variables include species (`r unique(penguins$species)`), island (`r unique(penguins$island)`), bill length (mm), bill depth (mm), flipper length (mm), body mass (g), sex, and year (year data was collected).

The mean flipper length of the penguins is `r round(mean(penguins$flipper_length_mm, na.rm = TRUE), 2)` mm.

```{r penguins_scatterplot}

## Scatterplot of Flipper Length vs Bill Length in Penguins

ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) +
  geom_point() +
  labs(title = "Flipper Length vs Bill Length in Penguins",
       x = "Bill Length (mm)",
       y = "Flipper Length (mm)",
       color = "Species")

## Save scatterplot

ggsave("penguin_scatter.png", width = 8, height = 6)

```

## Problem 2

```{r problem_2}

## Creating dataframe with numeric, logical and character vectors

problem2_df = tibble(
  norm_samp = rnorm(10),
  vec_logical = norm_samp > 0,
  vec_char = letters[1:10],
  vec_factor = factor(c("Level1", "Level2", "Level3", "Level1", "Level2", "Level3", "Level1", "Level2", "Level3", "Level1"))
)

```

Attempting to take the mean of each variable in problem2_df dataframe.

```{r problem_2_dfmean, error=TRUE}

## Mean of numeric vector
mean_numeric = mean(pull(problem2_df, norm_samp))

## Mean of logical vector
mean_logical = mean(pull(problem2_df, vec_logical))

## Mean of character vector
mean_char = mean(pull(problem2_df, vec_char))

## Mean of factor vector
mean_factor = mean(pull(problem2_df, vec_factor))

```

We can see that the numeric and logical vectors return numeric values. In the case of the logical vector from problem2_df, this is the proportion of TRUE values as R computes TRUE as 1 and FALSE as 0.

vec_char produces an error message as it is not possible for R to calculate the mean of character values. Similarly, attempts to calculate the average directly for vec_factor does not work as factors are categorical.
